The Splashback Radius of Planck-SZ clusters

### 2018-11-15 Paper_new/RSP_paper_new.tex
more changes

\pdfoutput=1
\documentclass[iop, apjl, twocolappendix, numberedappendix]{emulateapj}
%\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc}
%\usepackage{ae,aecompl}
\usepackage{float}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{txfonts}
\usepackage{ulem}
\usepackage{xspace}
\usepackage[colorlinks=true,urlcolor=blue,linkcolor=blue,citecolor=blue]{hyperref}
%\usepackage{nccmath}
\usepackage[utf8]{inputenc}
\newcommand{\fhat}[1]{\expandafter\hat#1}
\newcommand{\fbar}[1]{\expandafter\bar#1}
\newcommand\fnurl[2]{%
  \href{#2}{#1}\footnote{\url{#2}}%
}
\def\rs{r_{\rm s}}
\def\rhos{\rho_{\rm s}}
\def\se{s_{\rm e}}
\def\ftrans{f_{\rm trans}}
\def\rhog{\rho_{\rm g}}
\def\rhoginner{\rho_{\rm g}^{\rm inner}}
\def\rhogouter{\rho_{\rm g}^{\rm outer}}
\def\rt{r_{\rm t}}
\def\rout{r_{\rm out}}
\def\mpch{h^{-1}{\rm Mpc}}
\def\msunh{h^{-1}{\rm M_{\odot}}}
\shorttitle{Splashback radius of Planck-SZ clusters}
\shortauthors{Zürcher, More}

@@

\author{
Dominik Zürcher \altaffilmark{1, 2}
Surhud More \altaffilmark{3, 1}, 
}
\affil{
$^{1}$Kavli Institute for the Physics and Mathematics of the Universe (Kavli IPMU, WPI), 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8583, Japan \\ 
$^{2}$Department of Physics, ETH Zurich, Wolfgang Pauli Strasse 27, 8093 Zurich, Switzerland\\
$^{3}$The Inter-University Centre for Astronomy and Astrophysics, Post Bag 4, Ganeshkhind, Pune 411007, India
}

@@

%\begin{keywords}
%galaxies:clusters:general -- galaxies:halos -- methods:observational -- cosmology:dark matter -- cosmology:large-scale structure of Universe -- cosmology:observations
%\end{keywords}

@@

The final sample that we use consists of 596
galaxy clusters and is about an order of magnitude smaller compared
to the sample used in \citet{more2016detection}. The sample we use
in this paper has an average redshift of 0.177, and an average
cluster mass $M_{500c}$ of about $3.0 \times 10^{14}
h^{-1}$M$_{\sun}$. In Table~\ref{tab:cluster_catalogs} we compare
the main properties of the cluster catalog used in this work to that
used by \citet{more2016detection}. Additionaly, Figure~\ref{fig:placnk_summary}
visualizes the distribution of the masses and redshifts of the used clusters.
A map of the sky positions of the clusters
can be found in Figure~\ref{fig:planck_fig} in
Appendix~\ref{sec:figures}.

@@

\begin{figure*}
\centering{
  \includegraphics[scale=0.8]{derivatives_with_spline_no_mc.pdf}}
\caption{The distributions of the logarithmic derivatives of
the three-dimensional cross-correlation signals at the location of
the splashback radius are shown. In the left panel the distributions of the
derivatives of the full profiles are shown, whereas only the inner
halo term (namely $\rho_{\mathrm{in}}f_{\mathrm{trans}}$) is
shown in the right panel.}
   \label{fig:derivatives} 
\end{figure*}
\begin{figure*}
\centering{
    \includegraphics[scale=0.8]{splashback_with_spline_no_mc.pdf}}
\caption{The posterior distributions of the inferred
locations of steepest slope of the two- and three dimensional
cross-correlation signals. The distributions of the
locations of the two-dimensional steepening feature
($R_{\mathrm{sp}}^{\mathrm{2D}}$) are shown in the left panel, while the
distributions of the three-dimensional counterparts
($r_{\mathrm{sp}}^{\mathrm{3D}}$) are shown on the right. The
magnitude limit applied to the galaxy catalogs is indicated for each
distribution. Note that the colors of the distributions match with
the colors used in Figure~\ref{fig:2D_graphs} and
Figure~\ref{fig:3D_graphs}. For comparison, the location of the virial
radius $r_{\mathrm{200m}}$ as calculated from the average cluster
sample properties is indicated by the black, dotted lines.}
   \label{fig:splashback} 
\end{figure*}
\begin{figure}
    \includegraphics[scale=0.65]{color_separated_2D_with_spline_no_mc.pdf}
\caption{The two-dimensional cross-correlations
of the red and blue galaxy populations as inferred by
cross-correlating the cluster sample with the color-separated
subsamples that were extracted from the PS 21.5 galaxy catalog are
shown in the left panel, whereas the associated derivative profiles
are shown on the right. The vertical, shaded bands indicate the
locations of steepest slope of the cross-correlation signal as
inferred from the two subsamples, whereas the black, dashed lines
indicate the upper and lower bounds of the same feature but as
estimated from the full PS 21.5 galaxy catalog. }
   \label{fig:color_curve_2D} 
\end{figure}
\begin{figure*}
    \includegraphics[scale=0.76]{color_separated_3D_with_spline_no_mc.pdf}
\caption{We show the three-dimensional cross-correlations as
inferred by cross-correlating the cluster sample with the
color-separated galaxy subsamples, as well as the corresponding
splashback radii (indicated by the colored, vertical bands) and
their first and second order derivative profiles. The light-colored 
curves show the corresponding first and second order derivatives of the one-halo
term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:dk14}).
Here, the black, dashed lines indicate upper and lower bounds of the 
three-dimensional splashback radius corresponding to the full PS 21.5 galaxy sample.}
   \label{fig:color_curve_3D} 
\end{figure*}
%\begin{figure}
%    \includegraphics[width= \columnwidth]{color_fraction_with_spline_no_mc.pdf}
%\caption{Fractional contributions of the red and blue galaxy
%populations to the total galaxy number density. We show the
%locations of the splashback radii associated with the two subsamples
%with the colored, vertical bands and the upper and lower bounds of
%the splashback radius obtained from the full PS 21.5 catalog with
%the black, dashed lines.}
%   \label{fig:color_fraction} 
%\end{figure}
The right panels in Figure~\ref{fig:2D_graphs} show the
corresponding analytical derivatives of the two-dimensional
cross-correlations and the 68\% confidence interval based on our
model fits. The logarithmic derivatives show a distinct steepening
feature at around $1.3~\mpch$. The figure shows that the
location of the steepest slope does not change appreciably when the
magnitude limit is changed, even though we use a sample which is one
magnitude fainter than $M_{\rm i}-5\log h=-19.44$. The 68\%
confidence interval of the location of the steepest slope is
indicated by the vertical, shaded region. We also show the location
of the virial radius $r_{\mathrm{200m}}$ based on the Planck SZ mass estimate 
as a black, dotted line in each panel.

@@

% Star forming galaxies in models
% which quench their star formation on a timescale shorter than the
% orbital time scale are not expected to show presence of the splashback
% feature. A quantitative exploration of the implications is beyond the
% scope of this work, but will be followed up in the future.

% In Figure~\ref{fig:color_fraction}, we show the fractional
% contributions of the red and blue galaxies to the total,
% three-dimensional cross-correlation \footnote{The fractions are
% obtained by performing an appropriate number density weighting.}.
% The red galaxies dominate the central parts of the galaxy clusters
% but become less dominant as we move further away from the cluster center.
% We see an increase of the fractional contribution of
% red galaxies below the splashback radius. On scales larger than the
% splashback radius, the fractional contributions of the two populations become
% very similar.

@@

% As it can be seen from Table~\ref{tab:snr}, the signal-to-noise
% ratio of our measurement is lower when using the PS 22 catalog
% compared to the other two estimates. This has two causes: Firstly,
% the sky-coverage of the PS 22 catalog is $\approx$ 25\% lower than
% for PS 21 and PS 21.5 since more regions had to be excluded due to
% the survey depth being too low in those regions. Therefore, some
% clusters from the PSZ2 catalog are missed when the PS 22 catalog is
% exploited, which reduces the statistics. Secondly, as we consider
% fainter galaxies, the contamination from background galaxies
% increases. This requires a better estimate of the uncorrelated
% background component. A more precise estimation of the uncorrelated
% background contamination would be possible by using a random galaxy
% catalog in addition to the random cluster sample. This would allow
% to use the more sophisticated Landy \& Szalay estimator
% \citep{landy1993bias} to perform the background subtraction.

% Another approach to increase the statistics of the measurements would 
% be to increase the redshift range in which clusters are being considered. 
% We chose $0.03 \leq z \leq 0.33$ in order to keep the analysis comparable 
% to \citet{more2016detection}. However, this cut reduces the cluster sample 
% size. Increasing the redshift range has its own limitations, though. 
% It would be required to use a brighter magnitude limit for galaxies, 
% which reduces the number of galaxies that can be used to infer the 
% cross-correlation signal. This results in a trade-off which could be investigated
% to find the optimal redshift range. Alternatively, other cluster samples that do 
% not rely on optical cluster finding such as Xray based surveys might 
% serve as independent and larger cluster samples. 
% Another plausible approach would be to consider a smaller region of
% the sky, resulting in less clusters but with a deeper field of view,
% such as in the case of the Hyper Suprime-Cam at Subaru Telescope,
% which will survey $\approx 1400$ deg$^2$ of the sky and achieve a
% depth of 26.2 magnitudes \citep{takada2010subaru}.

The separation of the blue and red galaxy populations as described in
Section~\ref{sec:Color} is prone to photometric errors as explored in
the Appendix~\ref{sec:Errors}. A deeper galaxy catalog would be required in order to
confirm or rule out the existence of the splashback feature in the
blue galaxy population. The ongoing deep galaxy surveys such as the
Hyper Suprime-Cam (HSC) and the Dark Energy Survey (DES) would be able
to provide such data sets, albeit in a limited area. The Large
Synoptic Survey telescope (LSST) would eventually provide deep as well
as wide galaxy catalogs to eventually establish the locations of the
splashback radius at high significance.

@@

\begin{figure*}
\centering{
    \includegraphics[width= \textwidth]{color_vs_z.pdf}
\caption{Scatter plot showing the $(g_{\mathrm{P1}}-r_{\mathrm{P1}})$ colors
of the matched galaxies versus their redshifts, where the red dots correspond
to galaxies identified as red from the SDSS spectroscopic colors and blue dots
correspond to galaxies identified as blue, respectively. The black, solid line
indicates the spline cut separating the two populations and excluding the red 
galaxies from the blue population at a confidence level of $3 \sigma$.}
   \label{fig:g-r_vs_z} 
\end{figure*}

>>>>

\pdfoutput=1
\documentclass[iop, apjl, twocolappendix, numberedappendix]{emulateapj}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{txfonts}
\usepackage{ulem}
\usepackage{xspace}
\usepackage[colorlinks=true,urlcolor=blue,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage[utf8]{inputenc}
\newcommand{\fhat}[1]{\expandafter\hat#1}
\newcommand{\fbar}[1]{\expandafter\bar#1}
\newcommand\fnurl[2]{%
  \href{#2}{#1}\footnote{\url{#2}}%
}
\def\rs{r_{\rm s}}
\def\rhos{\rho_{\rm s}}
\def\se{s_{\rm e}}
\def\ftrans{f_{\rm trans}}
\def\rhog{\rho_{\rm g}}
\def\rhoginner{\rho_{\rm g}^{\rm inner}}
\def\rhogouter{\rho_{\rm g}^{\rm outer}}
\def\rt{r_{\rm t}}
\def\rout{r_{\rm out}}
\def\mpch{h^{-1}{\rm Mpc}}
\def\msunh{h^{-1}{\rm M_{\odot}}}
\shorttitle{Splashback radius of Planck-SZ clusters}
\shortauthors{Zürcher, More}

@@

\thanks{This work was submitted and defended as a Masters thesis by Dominik
Zuercher at the DPHYS in ETH Zurich in June 2018.}

\author{
Dominik Zürcher \altaffilmark{1, 2, 4}
Surhud More \altaffilmark{3, 1}, 
}
\affil{
$^{1}$Kavli Institute for the Physics and Mathematics of the Universe (Kavli IPMU, WPI), 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8583, Japan \\ 
$^{2}$Department of Physics, ETH Zurich, Wolfgang Pauli Strasse 27, 8093 Zurich, Switzerland\\
$^{3}$The Inter-University Centre for Astronomy and Astrophysics, Post Bag 4, Ganeshkhind, Pune 411007, India\\
$^{4}$ dominikz@phys.ethz.ch
}

@@

% \begin{keywords}
% galaxies:clusters:general -- galaxies:halos -- methods:observational -- cosmology:dark matter -- cosmology:large-scale structure of Universe -- cosmology:observations
% \end{keywords}

@@

The final sample that we use consists of 596
galaxy clusters and is about an order of magnitude smaller compared
to the sample used in \citet{more2016detection}. The sample we use
in this paper has an average redshift of 0.177, and an average
cluster mass $M_{500c}$ of about $3.0 \times 10^{14}
h^{-1}$M$_{\sun}$. In Table~\ref{tab:cluster_catalogs} we compare
the main properties of the cluster catalog used in this work to that
used by \citet{more2016detection}. Additionaly, Figure~\ref{fig:planck_summary}
visualizes the distribution of the masses and redshifts of the used clusters.
A map of the sky positions of the clusters
can be found in Figure~\ref{fig:planck_fig} in
Appendix~\ref{sec:figures}.

@@

\begin{figure*}
\centering{
  \includegraphics[scale=0.8]{derivatives_with_spline_no_mc.pdf}}
\caption{The distributions of the logarithmic derivatives of
the three-dimensional cross-correlation signals at the location of
the splashback radius are shown. In the left panel the distributions of the
derivatives of the full profiles are shown, whereas only the inner
halo term (namely $\rho_{\mathrm{in}}f_{\mathrm{trans}}$) is
shown in the right panel.}
   \label{fig:derivatives} 
\end{figure*}
\begin{figure*}
\centering{
    \includegraphics[scale=0.8]{splashback_with_spline_no_mc.pdf}}
\caption{The posterior distributions of the inferred
locations of steepest slope of the two- and three dimensional
cross-correlation signals. The distributions of the
locations of the two-dimensional steepening feature
($R_{\mathrm{sp}}^{\mathrm{2D}}$) are shown in the left panel, while the
distributions of the three-dimensional counterparts
($r_{\mathrm{sp}}^{\mathrm{3D}}$) are shown on the right. The
magnitude limit applied to the galaxy catalogs is indicated for each
distribution. Note that the colors of the distributions match with
the colors used in Figure~\ref{fig:2D_graphs} and
Figure~\ref{fig:3D_graphs}. For comparison, the location of the virial
radius $r_{\mathrm{200m}}$ as calculated from the average cluster
sample properties is indicated by the black, dotted lines.}
   \label{fig:splashback} 
\end{figure*}
\begin{figure}
    \includegraphics[scale=0.65]{color_separated_2D_with_spline_no_mc.pdf}
\caption{The two-dimensional cross-correlations
of the red and blue galaxy populations as inferred by
cross-correlating the cluster sample with the color-separated
subsamples that were extracted from the PS 21.5 galaxy catalog are
shown in the left panel, whereas the associated derivative profiles
are shown on the right. The vertical, shaded bands indicate the
locations of steepest slope of the cross-correlation signal as
inferred from the two subsamples, whereas the black, dashed lines
indicate the upper and lower bounds of the same feature but as
estimated from the full PS 21.5 galaxy catalog. }
   \label{fig:color_curve_2D} 
\end{figure}
\begin{figure*}
    \includegraphics[scale=0.76]{color_separated_3D_with_spline_no_mc.pdf}
\caption{We show the three-dimensional cross-correlations as
inferred by cross-correlating the cluster sample with the
color-separated galaxy subsamples, as well as the corresponding
splashback radii (indicated by the colored, vertical bands) and
their first and second order derivative profiles. The light-colored 
curves show the corresponding first and second order derivatives of the one-halo
term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:dk14}).
Here, the black, dashed lines indicate upper and lower bounds of the 
three-dimensional splashback radius corresponding to the full PS 21.5 galaxy sample.}
   \label{fig:color_curve_3D} 
\end{figure*}
The right panels in Figure~\ref{fig:2D_graphs} show the
corresponding analytical derivatives of the two-dimensional
cross-correlations and the 68\% confidence interval based on our
model fits. The logarithmic derivatives show a distinct steepening
feature at around $1.3~\mpch$. The figure shows that the
location of the steepest slope does not change appreciably when the
magnitude limit is changed, even though we use a sample which is one
magnitude fainter than $M_{\rm i}-5\log h=-19.44$. The 68\%
confidence interval of the location of the steepest slope is
indicated by the vertical, shaded region. We also show the location
of the virial radius $r_{\mathrm{200m}}$ based on the Planck SZ mass estimate 
as a black, dotted line in each panel.

@@

@@

The separation of the blue and red galaxy populations as described in
Section~\ref{sec:Color} is prone to photometric errors as explored in
the Appendix~\ref{sec:Errors}. A deeper galaxy catalog would be required in order to
confirm or rule out the existence of the splashback feature in the
blue galaxy population. The ongoing deep galaxy surveys such as the
Hyper Suprime-Cam \citep{Aihara2018} and the Dark Energy Survey \citep{Abbott2018} would be able
to provide such data sets, albeit in a limited area. The Large
Synoptic Survey telescope \citep{LSST2009} would eventually provide deep as well
as wide galaxy catalogs to eventually establish the locations of the
splashback radius at high significance.

@@

\begin{figure*}
\centering{
    \includegraphics[width= \textwidth]{color_vs_z.pdf}}
\caption{Scatter plot showing the $(g_{\mathrm{P1}}-r_{\mathrm{P1}})$ colors
of the matched galaxies versus their redshifts, where the red dots correspond
to galaxies identified as red from the SDSS spectroscopic colors and blue dots
correspond to galaxies identified as blue, respectively. The black, solid line
indicates the spline cut separating the two populations and excluding the red 
galaxies from the blue population at a confidence level of $3 \sigma$.}
   \label{fig:g-r_vs_z} 
\end{figure*}
